Concurrent Programming with CXL

### 2020-06-06 book/paper.tex
Added Italian Philosophers

\begin{quote}
\begin{itemize}
\item Do you see what the problem is?
\item Does it depend on \texttt{N}, the number of philosophers?
\item Can you come up with a solution?
\end{itemize}
\end{quote}

@@

@@

So far we have looked at how to protect a single resource.
Some types of resources may have multiple units or slots, each of which
If there are only $n$ units, no more than $n$ units can be used at a time;
if all are in use and a process comes along that needs one of the units,
it has to wait until another process releases one of the units.
Note that we cannot solve this problem simply using a lock per unit---allocating
a unit requires access to the entire collection.

A good example is the so-called \emph{bounded buffer}.  It is essentially
a queue implemented using a circular buffer of a certain length and two pointers:
one where items are inserted and one from where items are extracted.  If the
buffer is full, processes that want to add more items have to wait.  In the usual
formulation, processes that want to extract an item when the buffer is empty
also have to wait.
This problem is known as the ``Producer/Consumer Problem.''

@@

It would be easy to solve using a busy waiting
approach, but again, we would like to avoid busy waiting.
Introduced by the Dijkstra,
a \emph{semaphore} is a primitive that fits the bill.  A semaphore is essentially
a counter that can be incremented and decremented but is not allowed to go
below zero.  The semaphore counter is typically initialized to the number of
units of a resource initially available.
When allocating a resource, a process decrements the
counter using the \texttt{procure} or simply \texttt{P} operation.  The
\texttt{procure} operation blocks the invoking process if the counter is zero.
To release the resource, a process increments the resource using the
\texttt{vacate} or {V} operation.
Figure~\ref{fig:semaphore} shows the \texttt{synch} module implementation of
semaphores.

One can think of a semaphore as a generalization of a lock.  In fact, a
lock can be implemented by a semaphore initialized to 1.  Then \texttt{P}
acquires the lock, and \texttt{V} releases the lock.

>>>>

\begin{quote}
\begin{itemize}
\item Do you see what the problem is?
\item Does it depend on \texttt{N}, the number of philosophers?
\item Can you come up with a solution?
\item Does it matter in what order the philosophers lay down their forks?
\end{itemize}
\end{quote}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{../DinersSema.cxl}
\end{code}
\caption{Italian Philosophers.}
\label{fig:dinerssema}
\end{figure}

So far we have looked at how to protect a single resource.
Some types of resources may have multiple units.
If there are only $n$ units, no more than $n$ units can be used at a time;
if all are in use and a process comes along that needs one of the units,
it has to wait until another process releases one of the units.
Note that we cannot solve this problem simply using a lock per unit---allocating
a unit requires access to the entire collection.

Consider a variant of the Dining Philosophers (which were clearly Dutch---who
else need two forks to eat a meal?) that I call Italian Philosophers (who only
need one fork each to eat spaghetti).  Again, there are five philosophers.
This time, however, there is a glass sitting in the middle of the table with only
three forks in it.  A philosopher needs one of those forks to eat, but clearly,
no more than three philosophers can eat at a time.

\begin{quote}
\begin{itemize}
\item See if you can come up with a deadlock-free solution using locks.  Represent
    each fork as a lock.
\end{itemize}
\end{quote}

It's not easy.  It would be easy to come up with a solution that uses a single lock
and busy waiting, but we would like to avoid busy waiting.
Introduced by Dijkstra,
a \emph{semaphore} is a primitive that fits the bill.  A semaphore is essentially
a counter that can be incremented and decremented but is not allowed to go
below zero.  The semaphore counter is typically initialized to the number of
units of a resource initially available.
When allocating a resource, a process decrements the
counter using the \texttt{procure} or simply \texttt{P} operation.  The
\texttt{procure} operation blocks the invoking process if the counter is zero.
To release the resource, a process increments the resource using the
\texttt{vacate} or {V} operation.

Figure~\ref{fig:semaphore} shows the \texttt{synch} module implementation of
semaphores.
Figure~\ref{fig:dinerssema} shows a solution to the Italian Philosphers problem
using a semaphore.
One can think of a semaphore as a generalization of a lock.  In fact, a
lock can be implemented by a semaphore initialized to 1.  Then \texttt{P}
acquires the lock, and \texttt{V} releases the lock.

\chapter{Bounded Buffer}

@@

A good example of a resource with multiple units is the so-called \emph{bounded buffer}.
It is essentially
a queue implemented using a circular buffer of a certain length and two pointers:
one where items are inserted and one from where items are extracted.  If the
buffer is full, processes that want to add more items have to wait.  In the usual
formulation, processes that want to extract an item when the buffer is empty
also have to wait.
This problem is known as the ``Producer/Consumer Problem.''

@@
